Bootstrapping Multi-atlas Hippocampal Segmentation with \mb

### 2013-05-03 paper.Rnw
Methods: adni xval experiment now contains all registration and label fusion details; the useless histogram chart is gone

Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
of images and manual hippocampal segmentations from ADNI1 dataset.  This form of
cross-validation allows us to rigorously validate a large number of parameter
settings of \mb.

@@

Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
authors from the ADNI1 database (adni.loni.ucla.edu) between March 2012 and
August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
standardized dataset available from ADNI \footnote{
\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
\citep{Wyman2012}. This image collection contains uniformly preprocessed images
which have been designated to be the "best" after quality control.  All images
were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
Systems or Siemens Medical Solutions) at multiple sites using the protocol
described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR
= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
x 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.

@@

In this experiment, our dataset consists of 69 1.5T images and corresponding
manual segmentations taken from the ADNI1 project (see below for details). In each
validation round, the training set is used as an atlas library for both \mb and
basic multi-atlas. The remaining images are segmented by both methods, and the
resulting segmentations are compared to the manual segmentations for the images.

@@

\mb produces hippocampus segmentation volumes that are highly correlated with 
manual segmentation volumes (Figure \ref{fig:SZ-volumes}).  Additionally, as we 
saw in the previous experiments, \mb volumes show a fixed bias towards smaller 
volumes (although the bias is negligble), and a proportional bias towards 
producing smaller segmentations for larger hippocampi (Figure \ref{fig:FEP-SZ-Bland-Altman}.

@@

\begin{figure}
<<ADN1-FEP-Bland-Atlman, cache=TRUE, dependson='FEP-volumes', fig.width=4, fig.height=4>>=
bland_altman_plot <- function(x, y) {
  data = data.frame(x=x,y=y,diff=x-y)
  upper = mean(data$diff) + 2 * sd(data$diff)
  lower = mean(data$diff) - 2 * sd(data$diff)
  return(ggplot(data, aes(x= x, y = diff)) + 
    geom_hline(yintercept = mean(data$diff) + c(-2, 0, 2) * sd(data$diff), 
               linetype=2, color='brown') + 
    annotate("text", y = upper, x=Inf, 
              label=paste("std =",round(upper,0)), 
              hjust=1.1,vjust=1.5) + 
    annotate("text", y = lower, x=Inf, 
              label=paste("-std =",round(lower,0)), 
              hjust=1.1,vjust=1.5) + 
    geom_point())
}

bland_altman_plot(casted$Manual, casted$MAGeT) + 
  xlab('Mean manual and MAGeT Brain total Hippocampus volume') + 
  ylab('Manual - MAGeT Brain volume') + 
  geom_smooth(method="lm", formula = y~x)
@
  \caption{Bland-Altman plot comparing manual and \mb hippocampal volumes 
  when using five Winterburn atlases, and a 21 image template library.}
  \label{fig:FEP-SZ-Bland-Altman}
\end{figure}

@@

% paired t-test against SNT to show similarity (should be low) 
% Kruskal-Willis against  (should be high?)
% post-hoc power analysis to say we can detect this change
\begin{figure}[h]
<<ADNI-scr-histogram-prep,cache=T>>=
melted   = melt(totals.complete, measure.vars=c( "SNT","FSL","FS","MAPER","MAGeT"))
binwidth = 250

# prepare the normals
grid <- with(melted, seq(min(value), max(value), length = 100))
normaldf <- ddply(melted, c("variable", "DX"), function(df) {
  data.frame( 
    value = grid,
    density = dnorm(grid, mean(df$value), sd(df$value)),
    freq = length(df$value) * dnorm(grid, mean(df$value), sd(df$value)) * binwidth,
    length = length(df$value)
  )
})

testsdf <- ddply(melted, c("variable", "DX"), function(df) {
  shapiro = shapiro.test(df$value)
  data.frame(
    shapiro.w = shapiro$statistic,
    shapiro.p = shapiro$p.value,
    mean = mean(df$value),
    sd = sd(df$value)
  )
})

melted.by.snt = melt(totals.complete, measure.vars=c("MAGeT","FSL", "FS", "MAPER"))
snt.tests <- ddply(melted.by.snt, c("variable", "DX"), function(df) {
  ks = ks.test(df$SNT, df$value)
  data.frame( 
    ks.D = ks$statistic,
    ks.p = ks$p.value
  )
}) 
@

<<ADNI1-scr-histogram,dependson='ADNI-scr-histogram-prep',cache=T>>=

ggplot(data=melted, aes(x=value)) + 
  geom_histogram(binwidth=binwidth, alpha=0.8) + 
  geom_line(aes(y=freq, colour='red'), data=normaldf) +
  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=1.5,size=1, 
                label=paste("list(W ==",signif(shapiro.w, digits=3), ", p == ", 
                            round(shapiro.p,digits=3), ")")), 
            parse=TRUE, data=testsdf) +
  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=3,size=1, 
                label=paste("list(D ==",signif(ks.D, digits=3), ", p == ", 
                            round(ks.p,digits=3), ")")), 
            parse=TRUE, data=snt.tests) + 
  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=5.5,size=1, 
                label=paste("sigma ==",signif(sd, digits=3))), 
            parse=TRUE, data=testsdf) + 
  geom_text(aes(y=Inf,x=Inf,hjust=1.1,vjust=6,size=1, 
                label=paste("mu ==",signif(mean, digits=3))), 
            parse=TRUE,data=testsdf) + 
  #xlim(0, 10000) + 
  ylim(0,40) + 
  facet_grid(variable ~ DX) + 
  scale_colour_discrete(guide="none") + 
  scale_size_continuous(guide="none") + 
  labs(x=expression(paste("Total hippocampal volume (", mm^3, ")")),
       y="Frequency") + 
  theme(axis.title.x = element_text(vjust=-0.1))
@
  \caption{{\bf ADNI Baseline cohort.} Comparison of total hippocampal volumes
  as measured by SNT, FSL, FreeSurfer (FS), MAPER, and \mb (MAGeT). A
  fitted normal curve is shown in red. $W$ is the Shapiro-Wilkes test statistic
  measuring normality of the data (signficance indicates non-normality). $D$ is
  the Kolmogorov-Smirnov test statistic measuring the goodness-of-fit between
  the distribution of measured volumes and SNT volumes (significance indicates a
  difference).}
  \label{fig:ADNI-scr-histogram}
\end{figure}

\begin{figure}
  \includegraphics[width=6in,height=3in]{figure/ADNI1_MAGeT_montage/figure.pdf}
  \caption{A caption}
  \label{fig:ADNI1-segmentation-montage}
\end{figure}

>>>>

Monte Carlo Cross-Validation (MCCV) \citep{Shao1993} was performed using a pool
of images and manual hippocampal segmentations from ADNI1 dataset.  This form of
cross-validation allows us to rigorously validate a large number of parameter
settings of \mb (atlas and template library sizes, registration algorithm, and 
label fusion method) and select the best parameters to use in subsequent 
experiments.

@@

Clinical, demographic and pre-processed T1-weighted MRI were downloaded by the
authors from the ADNI database (adni.loni.ucla.edu) between March 2012 and
August 2012. The image dataset download was the "ADNI1:Complete 1Yr 1.5T"
standardized dataset available from ADNI \footnote{
\url{http://adni.loni.ucla.edu/methods/mri-analysis/adni-standardized-data/}}
\citep{Wyman2012}. This image collection contains uniformly preprocessed images
which have been designated to be the "best" after quality control.  All images
were acquired using 1.5T scanners (General Electric Healthcare, Philips Medical
Systems or Siemens Medical Solutions) at multiple sites using the protocol
described in \citep{Jack2008}.  Representative 1.5T imaging parameters were TR
= 2400ms, TI = 1000ms, TE = 3.5ms, flip angle = $8{\circ}$, field of view = 240
x 240mm, a $192 \times 192 \times 166$ matrix ($x$, $y$, and $z$ directions)
yielding a voxel resolution of $1.25 \times 1.25 \times 1.2mm^3$.

@@

In this experiment, our dataset consists of 69 1.5T images and corresponding
manual segmentations. In each validation round, the dataset is partitioned into 
a training set consisting of images and their manual labels to be used as an atlas 
library, and a validation set consisting of the remaining images segmented by both 
\mb and multi-atlas. The resulting segmentations are compared to the manual 
segmentations for the images.

@@

\mb produces hippocampus segmentation volumes that are highly correlated with 
manual segmentation volumes (Figure \ref{fig:SZ-volumes}).

@@

@@
